---
title: "Machine learning workflow using tidymodels"
subtitle: "PART1: BASIC USAGE"
author: "YYS"
date: "2024-10-5"
toc: true
number-sections: true
format:
  html:
    theme: litera
---

🌀 Workflow of Machine learning in R with package `tidymodels`

-   Strength of `tidymodels`

-   Basic usage

-   For workflow or iteration

------------------------------------------------------------------------

`tidymodels`, like `tidyverse`, have the same good character: "tidy"

> Good system can help build well ideas and acclerate our work.

As we all know, there exist lots of machine learning methods or packages in R, FROM different authors. So huge different usage due to so many parameters in each function.

Fortunately, `tidymodels` union these packages!

Not only use it for machine learning, but also in prepare data (including split data set into training and test, impute missing values, one-hot coding, normalize...etc.)

------------------------------------------------------------------------

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = F, message = F, dpi = 300)
```

```{r}
#| include: false
rm(list = ls())
```

------------------------------------------------------------------------

# load packages and data

The first thing, load packages needed.

```{r}
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(ggsci)
```

# basic usage

Here, we will use two examples to show how to do machine learning with tidymodels.

There are two kind of problems we will meet: **classification** and **regression**

> The model choice, depend on our interesting outcome (continous, category, or counting data).

------------------------------------------------------------------------

The first example, is the most used for classification question.

Let us using data named `penguins` fromo packages `palmerpenguins`, we want to predict `sex`(male and female) of each penguin.

```{r}
penguins |> glimpse()
```

Some EDA (explore data analysis) is necessary and interesting before we modelling.

```{r}
penguins |> count(sex)
```

```{r}
#| fig-width: 8
#| fig-height: 4
penguins |> 
  filter(!is.na(sex)) |> 
  ggplot(aes(flipper_length_mm, bill_length_mm, 
             color = sex, size = body_mass_g)) +
  geom_point(alpha = 0.7) +
  facet_wrap( ~ species)
```

There exist NA, so we drop them, and other 2 variables. `penguins_df` is the data we will use.

```{r}
penguins_df <- penguins |> 
  filter(!is.na(sex)) |> 
  select(-island) |> 
  rename(outcome = sex)
```

## modeling

1.  data split, the first step

-   Generally, we will divide a full dataset into two parts: training and testing set

-   then, in the training set, we will divide it again, using `bootstraps` or `cross-validation`, for model evaluate and high-parameter selection

    -   we can use `bootstraps` for resample with replacement or

    -   use `vfold_cv` for cross-validation

    -   FOR more detail of bootstraps and cross-validation, just `Google` them!

```{r}
set.seed(123)
df_split <- initial_split(penguins_df, strata = outcome)
df_train <- training(df_split)
df_test <- testing(df_split) 

set.seed(234)
df_folds_bs <- bootstraps(df_train, strata = outcome)
 

set.seed(345)
df_folds_cv <- vfold_cv(df_train, strata = outcome)
```

2.  build recipe

we can build formula in our model, and then some prepare such as imputing, dummy coding......

for more information, Google `step_ with recipe`

```{r}
df_recipe <- 
  recipe(formula = outcome ~ ., data = df_train) # |> 
  # step_impute_median(variable_1, variable_2, ...)|> 
  # step_dummy(variable_3, variable_4, ...)
```

3.  modelling

we use two model here: logistic and random forest.

we set engine such as `glm` and `ranger`, meanwhile, we add `importance = "impurity"` for further variable importance computation;

Then we add recipe above and model with engine, link them use `workflow()`;

next step, we bootstraps the model in training set, to see the performance.

```{r}
glm_spec <- logistic_reg() |> 
  set_engine("glm")

glm_wf <- workflow() |> 
  add_recipe(df_recipe) |> 
  add_model(glm_spec)

glm_rs <- glm_wf |> 
  fit_resamples(
    resamples = df_folds_bs,
    control = control_resamples(save_pred = T, verbose = T)
  )
glm_rs
```

```{r}
rf_spec <- rand_forest() |> 
  set_mode("classification") |> 
  set_engine("ranger", importance = "impurity")

rf_wf <- workflow() |> 
  add_recipe(df_recipe) |> 
  add_model(rf_spec)

rf_rs <- rf_wf |> 
  fit_resamples(
    resamples = df_folds_bs,
    control = control_resamples(save_pred = T, verbose = T)
  )
rf_rs
```

Now, we will find a tidy PHENOMENON, the logistic and random forest model are extremely same!

## evaluate

Next, we want to know how the performance of our model.

some metrics like: accuracy, roc_auc, sens, spec, recall, f_meas......

🌐[Metric types in yardstick](https://yardstick.tidymodels.org/articles/metric-types.html)

we can add more metrics into `fit_resamples` through `metrics = metric_set()`

### basic metrics

```{r}
collect_metrics(glm_rs)
```

also for `rf_rs`

```{r}
collect_metrics(rf_rs)
```

### confusion matrix

```{r}
glm_rs |> 
  conf_mat_resampled()
```

### ROC

```{r}
glm_rs |> 
  collect_predictions() |> 
  group_by(id) |> 
  roc_curve(outcome, .pred_female) |>   # 后面也可以直接 autoplot()
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = F, alpha = 0.6, size = 1.2) +
  coord_equal()
```

## in test final model

```{r}
glm_final <- glm_wf |> 
  last_fit(df_split)
glm_final
```

```{r}
rf_final <- rf_wf |> 
  last_fit(df_split)
rf_final
```

### metric in test

```{r}
collect_metrics(glm_final)
```

```{r}
collect_metrics(rf_final)
```

### pred probility

```{r}
collect_predictions(glm_final) |> 
  arrange(.pred_female) |> 
  head(10)
```

### ROC in test

```{r}
rbind(
  collect_predictions(glm_final) |>
    roc_curve(outcome, .pred_female) |>
    mutate(Model = "Logistic (AUC = 0.938)")
  ,
  collect_predictions(rf_final) |>
    roc_curve(outcome, .pred_female) |>
    mutate(Model = "Random forest (AUC = 0.933)")
) |>
  mutate(Model = fct_relevel(Model,
                             "Logistic (AUC = 0.938)",
                             "Random forest (AUC = 0.933)",
                             )) |> 
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_abline(lty = 2,
              color = "gray80",
              linewidth = 1.5) +
  geom_path(aes(color = Model), alpha = 0.8, linewidth = 1) +
  labs(x = "1 - Specificity", y = "Sensitivity") +
  coord_equal() +
  scale_color_lancet() +
  theme_classic() +
  theme(legend.position = c(0.75, 0.25))
```

### conf mat

```{r}
collect_predictions(glm_final) |> 
  conf_mat(outcome, .pred_class)
```

### estimate coef

```{r}
glm_final$.workflow[[1]] |> 
  tidy(exponetiate = T) |> 
  arrange(estimate)
```

------------------------------------------------------------------------

This is the first part of `tidymodels`;

🔜In next part, we will introduce how to select high-parameters

------------------------------------------------------------------------

😃Thank you to Julia Silge, an exceptionally talented data scientist. I have greatly enhanced my understanding of `tidymodels` through her informative YouTube channel.
